Secure Composition of Robust and Optimising Compilers

### 2023-07-10 popl/main.tex

This section defines programming languages that the secure compilers defined in the next section will use.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
However, not all $\Ltms$ programs are \gls{sms}.
That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.

This paper considers \gls{cct} as an architecture-dependent property that should not be one of the programmer's concerns~\cite{cauligi2019fact}.
Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
In spirit of this, the language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.

@@

$\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.
The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.
The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model whenever an event with a location is projected.
The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected, e.g., $\msfilterLtms{\src{\ev{Alloc}\ \loc\ n;\comp;\securitytag}}=\ev{Alloc}\ \delta(\loc)\ n;\comp;\securitytag$.
But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.

@@

$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
For allocated ($\trg{\poisonless}$) arrays bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
%
The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.

@@

While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode\footnote{As present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.}.
For the other languages seen earlier, this mode is intuitively always-on.
\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present $\Omega$ also in the shared lang defs
To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
If the register is turned on, no secrets can appear on traces.
If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access, which is taken care of by the projection to specification events.
\Cref{ex:lscct} illustrates these changes.

The projection to the specification trace model is mostly straightforward.
But for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value.
The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
With this technical setup, the information wether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.

@@

  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
  But, if it is turned on, then the branching event does not fire anymore.
  Moreover, writing to memory now yields the pre-event $\obj{\widehat{Set}}$, whose difference to $\obj{Set}$ becomes apparent when looking at the projection to the specification trace model.
  Projecting the above trace to the specification trace model yields the following trace (emitting empty events):
  $\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock\cdot\ev{Branch\ 0}\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock\cdot\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock\cdot\ev{Set}\ \loc_{y}\ 2;\sandboxtag;\unlock$.
  Clearly, secrets are hidden ($\unlock$) after this projection for the section of the program where the data independent timing mode is turned on, while still containing the appropriate events to check for \gls{ms}.
\end{example}

\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
\begin{figure}[!h]
  \begin{center}
    \begin{tikzpicture}
      \node (S) {$\src{L_{\tmssafe}}$};
      \node[right=1.5 of S] (T) {$\trg{L}$};
      \node[right=1.5 of T] (M) {$\irl{L_{\mssafe}}$};
      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
      \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
      \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};

@@

      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
    \end{tikzpicture}
  \end{center}
  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}. %
    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and strikethrough edges classic proof techniques. %
    The dashed lines partition the graph into the sections where the respective theorems are presented.
  }\label{fig:pipeline}
\end{figure}

This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
Since these properties compose into \gls{ms}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
The last step of the chain ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
The final result is that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
  \begin{align*}
%    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
                                                                                                \left[\cca(\src{\expr})\right] %
                                                                                                 }{\labort[\trg]}}
  \end{align*}

The compiler only inserts a dynamic typecheck to respect the well-typedness of function bodies.
Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
By inserting the check, the compiler ensures that execution does not proceed in such cases.
The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).

Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments are valid.

@@

\subsubsection{Proof Sketch of Robust Safety Property Preservation}

The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for similar theorems in this paper.

When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.
By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.
The proof now needs to relate the two execution with eachother.
\Cref{fig:proofdiag:rtp} visualizes this situation.
At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again. Recursive calls between context and component functions are disregarded.}.
Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}[state/.style={minimum height=0.6cm}]
      % relative horizontal/vertical distance between states
      \pgfmathsetmacro{\hdist}{0.95}
      \pgfmathsetmacro{\vdist}{1.35}
      \pgfmathsetmacro{\halfvdist}{0.725}

@@

@@

The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks to enforce \gls{sms} whenever reading from or writing to memory.
To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
Again, linking this back to the \texttt{strncpy} example (\Cref{ex:strncpy:sms}) corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
In turn, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
\MPin{
	example code?
    MK: still needed now that we reference ex. 3.10?
}

@@

\Cref{thm:ccab:rtp:ms} follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.

@@

\begin{center}
  $$
  \begin{array}{rcll}
    \cccf(\irl{\llet{x}{n}{\expr}}) & = & \cccf(\irl{\expr\subst{n}{x}}) &\\
    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = & \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}\\
  \end{array}
  $$
\end{center}
\MPin{
	substitution are a runtime semantics thing.
	what you want is a compiler with 2 arguments, where in one you keep track of the static substitions to perform,
	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
  %
  MK: No? I get how they are used as part of the runtime semantics, but a compiler can just as well make use of a substitution function that models the runtime semantics. I see how these two are different. However, I think it's easier in this case to be a bit inprecise, since adding another argument to the compiler breaks our compositionality framework (unless it's hidden with a monad...) so, IMO, we should view substitution here as a meta-level, but still language-dependent, function that the compiler has access to. If we don't do this, we essentially re-encode the definition of subst into the compiler
}

@@

\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
\end{theorem}
\MPin{
This you can write the proof of.
Probably even in coq
%
MK: Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}
}

>>>>

This section defines programming languages that the secure compilers defined in the next section will use.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
However, not all $\Ltms$ programs are \gls{sms}.
That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.

This paper considers \gls{cct} as an architecture-dependent property writing \gls{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
In spirit of this, language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.

@@

$\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.\MP{fabulous paper from amal?relevant?}
\MP{and therefore omitted?}
The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms} (\Cref{thm:wt:tms}).
The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model whenever an event with a location is projected.
The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected, e.g., $\msfilterLtms{\src{\ev{Alloc}\ \loc\ n;\comp;\securitytag}}=\ev{Alloc}\ \delta(\loc)\ n;\comp;\securitytag$.
But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.
\MPin{ok let's add details}

@@

$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
For allocated ($\trg{\poisonless}$) arrays bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
%
The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
\MP{ok, details}

@@

While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
For the other languages seen earlier, this mode is intuitively always-on.
\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present $\Omega$ also in the shared lang defs
To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
If the register is turned on, no secrets can appear on traces.
If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access, which is taken care of by the projection to specification events.
\Cref{ex:lscct} illustrates these changes.
\MPin{formalisation additions in algin* env}

The projection to the specification trace model is mostly straightforward.
But for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value.
The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
With this technical setup, the information wether\MP{spelling} a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.

@@

  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
  But, if it is turned on, then the branching event does not fire anymore.
  Moreover, writing to memory now yields the pre-event $\obj{\widehat{Set}}$, whose difference to $\obj{Set}$ becomes apparent when looking at the projection to the specification trace model.
  \MPin{shouldn't we have 2 traces, one with the bit set, one without}
  Projecting the above trace to the specification trace model yields the following trace (emitting empty events):
  $\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock\cdot\ev{Branch\ 0}\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock\cdot\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock\cdot\ev{Set}\ \loc_{y}\ 2;\sandboxtag;\unlock$.
  Clearly, secrets are hidden ($\unlock$) after this projection for the section of the program where the data independent timing mode is turned on, while still containing the appropriate events to check for \gls{ms}.
\end{example}

\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
\begin{figure}[!h]
  \begin{center}
    \begin{tikzpicture}
      \node (S) {$\src{L_{\tmssafe}}$};
      \node[right=1.5 of S] (T) {$\trg{L}$};
      \node[right=1.5 of T] (M) {$\irl{L_{\mssafe}}$};
      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
      \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
      \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};

@@

      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
    \end{tikzpicture}
  \end{center}
  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}. %
    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and strikethrough edges classic proof techniques. %
    The dashed lines partition the graph into the sections where the respective theorems are presented.
  }\label{fig:pipeline}
\end{figure}
The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
Since these properties compose into \gls{ms}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
The last step of the chain ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
The final result is that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
\MPin{some text should introduce the compiler defs}

  \begin{align*}
%    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
                                                                                                \left[\cca(\src{\expr})\right] %
                                                                                                 }{\labort[\trg]}}
  \end{align*}

The first compiler only inserts a dynamic typecheck to respect the well-typedness of function bodies.
Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
By inserting the check, the compiler ensures that execution does not proceed in such cases.
The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).

Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.

@@

\subsubsection{Proving Robust Safety Property Preservation}

The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for the following theorems of this section.
\MP{so what?}

When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
\MP{passive + not justified}
The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
\MP{no, the goal is to BT}
To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.\MP{passive}
By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.\MP{passive passive passive, and imprecise}
The proof now needs to relate the two execution with each other.
\Cref{fig:proofdiag:rtp} visualizes this situation\MP{imprec}.
\begin{figure}[!ht]
  \begin{center}
    \begin{tikzpicture}[state/.style={minimum height=0.6cm}]
      % relative horizontal/vertical distance between states
      \pgfmathsetmacro{\hdist}{0.95}
      \pgfmathsetmacro{\vdist}{1.35}
      \pgfmathsetmacro{\halfvdist}{0.725}

@@

\MPin{
	thicker lines and dots needed, not visible otherwise

At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again. Recursive calls between context and component functions are disregarded.}.
Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.

@@

The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks to enforce \gls{sms} whenever reading from or writing to memory.
To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
Again, linking this back to the \texttt{strncpy} example (\Cref{ex:strncpy:sms}) corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
In turn, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
\MPin{
	show the result of the compilation as listing, and then describe the trace a bit more in detail ?
}

@@

\Cref{thm:ccab:rtp:ms}\MP{states ... } follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.

@@

\begin{center}
  $$
  \begin{array}{rcll}
    \cccf(\irl{\llet{x}{n}{\expr}}) & = & \cccf(\irl{\expr\subst{n}{x}}) &\\
    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = & \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}\\
  \end{array}
  $$
\end{center}
\MPin{
	substitution are a runtime semantics thing.
	what you want is a compiler with 2 arguments, where in one you keep track of the static substitions to perform,
	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
  %
  \MKin{
  MK: No? I get how they are used as part of the runtime semantics, but a compiler can just as well make use of a substitution function that models the runtime semantics. I see how these two are different. However, I think it's easier in this case to be a bit inprecise, since adding another argument to the compiler breaks our compositionality framework (unless it's hidden with a monad...) so, IMO, we should view substitution here as a meta-level, but still language-dependent, function that the compiler has access to. If we don't do this, we essentially re-encode the definition of subst into the compiler
  }
}

@@

\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
\end{theorem}
\MPin{
This you can write the proof of.
Probably even in coq
%
	\MKin{ 
	Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}

	\MPin{yes it should}}
}
